---
title: "Boosted HP Filter"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r setup}
library(bHP)
```

## HP filter

The Hodrick-Prescott filter (HP filter; Hodrick and Prescott, 1997) is one of the fundamental statistical 
tools. Given time series data $(x_{t})_{t=1}^n$
the HP method decomposes the series into
two additive components --- a trend component $f_{t}$
and a residual or cyclical component $c_{t}$, estimated as
\[
(\hat{f}_{t}^{HP} )
 =\arg\min_{ (f_{t} )}  \{ \sum_{t=1}^{n} (x_{t}-f_{t} )^{2}
 +\lambda\sum_{t=2}^{n} (\Delta^ 2 f_{t}  )^{2} \},
\]
 and
\[
(\hat{c}_{t}^{HP} )=( x_t-\hat{f}_{t}^{HP}) 
\]
 where $\Delta f_{t}=f_{t}-f_{t-1}$,
and $\Delta^2 f_{t}= \Delta f_{t}- \Delta f_{t-1} = f_{t}- 2 f_{t-1} + f_{t-2}$,
and $\lambda\geq 0$
is a tuning parameter that controls the extent of the penalty. The choice of the tuning parameter is crucial for the behavior of the HP filter. 

If $D'$ is the rectangular $\left(n-2\right)\times n$ matrix
with second differencing vector $d=\begin{pmatrix}1 & -2 & 1\end{pmatrix}'$ along
the leading tri-diagonals and  $I_{n}$ is the $n\times n$ identity
matrix, the explicit form of the trend solution is
\begin{equation} \label{tt1}
\widehat{f}^{\mathrm{HP}}=Sx,
\end{equation}
where $S=\left(I_n+\lambda DD'\right)^{-1}$ is a deterministic operator and $x=(x_1,...,x_n)'$ is the sample data.
The smoothed component $\widehat{f}^{\mathrm{HP}}$ is interpreted as the estimated 
trend and 
\[
\widehat{c}^{\mathrm{HP}}=x-\widehat{f}^{\mathrm{HP}}=\left(I_{n}-S\right)x
\]
as the estimated cyclical or stationary component. 

In practice, Hodrick and Prescott (1997) recommend $\lambda=1600$ for quarterly data, and this number and its sampling frequency adjusted version (Ravn and Uhlig, year) are widely adopted in applications. However, recent research (Phillips and Jin, 2015; Hamilton 2018) find the "gold standard" is unsatisfactory for the length of time series that often encountered in macroeconomic studies.


## Boosted HP filter

Phillips and Shi (2019) propose the boosted HP filter (bHP). The intuition is that, if the cyclical component $\widehat{c}_{t}^{\mathrm{HP}}$ still exhibits trending behavior after HP filtering, we continue to apply the HP filter to
$\widehat{c}^{\mathrm{HP}}$ to remove the leftover trend residual. After a second fitting, the cyclical component can be written as
\[
\widehat{c}^{\left(2\right)}=\left(I_{n}-S\right)\widehat{c}^{\mathrm{HP}}=\left(I_{n}-S\right)^{2}x,
\]
where the superscript ``$\left(2\right)$'' indicates that the HP
filter is fitted twice. The corresponding trend component becomes
\[
\widehat{f}^{\left(2\right)}=x-\widehat{c}^{\left(2\right)}=\left(I_{n}-\left(I_{n}-S\right)^{2}\right)x.
\]
If $\widehat{c}^{\left(2\right)}$ continues to exhibit trend behavior, the filtering process may be continued for a third or further time. 
After $m$ repeated applications of the filter, the cyclical and trend component are 
\[
	\widehat{c}^{\left(m\right)}  =  \left(I_{n}-S\right)\widehat{c}^{\left(m-1\right)}=\left(I_{n}-S\right)^{m}x \\
	\widehat{f}^{\left(m\right)}  =  x-\widehat{c}^{\left(m\right)}=B_{m}x, 
\]
where $B_{m}=I_{n}-\left(I_{n}-S\right)^{m}.$

The boosted HP filter introduces the number of iterations $m$ as an additional tuning parameter. In practice, it is recommended that we choose $\lambda$ according to the convention, say $\lambda = 1600$ for quarterly data, and then monitor a stopping criterion as the iteration proceeds. Phillips and Shi (2019) suggest using either the ADF test or the BIC as the stopping criterion. 

This package `bHP` automates the boosted HP filter. The main function is named `BoostedHP`. 
The user chooses the two tuning parameters `lambda` for $\lambda$ and `stopping` for the stopping criterion. In particular, `stopping` has three options:

* `BIC` for the BIC stopping criterion
* `adf` for the ADF stopping criterion (default $p$-value is 5\%)
* `nonstop` keeps iteration until it reaches `Max_iter` (default is 100 iterations).


## Examples

```{r}
lam <- 100 # tuning parameter for the annual data
data(IRE) # load the data 'IRE'

bx_HP <- BoostedHP(IRE, lambda = lam, iter= FALSE)$cycle
bx_BIC <- BoostedHP(IRE, lambda = lam, stopping = "BIC")$cycle
bx_ADF <- BoostedHP(IRE, lambda = lam, stopping = "adf")$cycle

bx_nonstop <- BoostedHP(IRE, lambda = lam, iter= TRUE, stopping = "nonstop") %>% residuals()

matplot( cbind(IRE, bx_HP, bx_BIC, bx_ADF, bx_nonstop))
```

```{r, eval=FALSE}
#-------- plot -----------

?plot.bHP

#--------- start to plot the content of bHP -----------------

#--------- for dynamic style (default)--------
plot(bx_ADF)

plot(bx_ADF, iteration_location = "upright") # change the location of text

plot(bx_ADF, iteration_location = c(30,12)) # assign the location of text by x-y co-ordinates

plot(bx_BIC, interval_t = 0.8 ) # change the time interval of animation

plot(bx_nonstop, cex_legend = 2, cex_text = 3) # change the magnification of legend and text

# change the color
plot(bx_ADF,main = "dynamic graph with new color",col_raw = "#685F74", col_trend_h = "#39A1A8", col_trend_f = "#DD4B4F", col_pvalue_BIC = "#E96145")

plot(bx_ADF,main = "dynamic graph with new trancparency setting",raw_alpha = 200, trend_h_alpha = 55, trend_f_alpha = 250, pvalue_BIC_alpha = 250)

plot(bx_HP)
# nonstop-iter' bHP doesn't have dynamic picture: returning NA

#--------- for JS style ----------

plot(bx_ADF,plot_type = "JS")

# change the color
plot(bx_ADF,plot_type = "JS",main = "Js graph with new color", col_raw = "#685F74", col_trend_f = "#DD4B4F", col_pvalue_BIC = "#39A1A8")

plot(bx_BIC,plot_type = "JS")

plot(bx_nonstop,plot_type = "JS")

plot(bx_HP,plot_type = "JS")

#--------- for static style ----------

plot(bx_ADF,plot_type = "static",cex_legend = 0.7, cex_text = 0.8 )

plot(bx_HP,plot_type = "static")

plot(bx_BIC,plot_type = "static",cex_legend = 0.7, cex_text = 0.8 )

plot(bx_nonstop,plot_type = "static",cex_legend = 0.8, cex_text = 0.8 )

#----------- print -------------------------------

?print.bHP

#--------- start to print the content of bHP -----------------
print(bx_ADF)

print(bx_ADF, Head = F, Tail = T, trend_hist = F)

print(bx_ADF, Head = T, Tail = T, trend_hist = F)

print(bx_ADF, Head = F, Tail = F, trend_hist = F)

print(bx_BIC, Head = F, Tail = F, trend_hist = T, select_trend_hist = 1:bx_BIC$iter_num)

print(bx_BIC, Head = F, Tail = F, trend_hist = T,  select_trend_hist = c(1,3,5))

# when the trend_hist is FALSE, select_trend_hist is invalid
print(bx_BIC, Head = F, Tail = F, trend_hist = F, select_trend_hist = c(1,3,5))

print(bx_BIC, Head = F, Tail = T, trend_hist = F, print_type = "latex")

print(bx_BIC, Head = F, Tail = T, trend_hist = F, print_type = "html")

# show the generic print function output
print(bx_ADF, type = "generic default")



#------------------ summary -----------------

?summary.bHP

summary(bx_ADF)
summary(bx_BIC)
summary(bx_nonstop)
summary(bx_HP)

#------------------ predict -----------------

?predict.bHP

predict(bx_HP) #Iterated number of HP filter: 1

predict(bx_ADF) #Iterated number of HP filter: 19

predict(bx_BIC) #Iterated number of HP filter: 5

predict(bx_nonstop) #Iterated number of HP filter: 99


#------------------ residuals -----------------

?residuals.bHP

residuals(bx_HP) #Iterated number of HP filter: 1

residuals(bx_ADF) #Iterated number of HP filter: 19

#------------------ BIC -------------------------

?BIC.bHP

BIC(bx_BIC)

#Retrun the value path of BIC.
#Iterated number of HP filter: 5
#Keep the path of BIC till iterated 6 times to show the tuning point.
#[1] 1.586255 1.366335 1.293931 1.264323 1.254397 1.254620

BIC(bx_nonstop)

#Retrun the BIC path of nonstop.
#Iterated number of HP filter: 99
#Keep the path of BIC till iterated 100 times to show the tuning point.
#[1] 1.586255 1.366335 1.293931 1.264323 1.254397 1.254620 1.260345 1.269139 1.279670 1.291179
#[11] 1.303223 ...


### If the test type is not "adf", Pvalue.bHP will return error

# raw HP filter
BIC(bx_HP)

# Error in BIC.bHP(bx_HP) :
# The stationary test type is nonstop-iter, not BIC or nonstop.


# by ADF
BIC(bx_ADF)

#Error in BIC.bHP(bx_ADF) :
#The stationary test type is adf, not BIC or nonstop.

```




